---
title: "CM515 Assignment 5"
author: "Kamal Neupane"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
```

# Instructions 

The aim of this assignment is to provide you with an opportunity to sharpen your skills in using ggplot(). While some of the plots you create may resemble those we covered in class, it is essential that they are original. 
 
 Rules for the Assignment:
 
 1.All data sets except iris are fair game. That includes past data sets used in the class, sets built into R, your own data, and even online data.
 
 2. All graphs must include axis labels, plot title, a theme of your choice, and a brief description/interpretation of the plot (2-3 sentences).


It's worth noting that ggplot() is a widely-used tool, and there are numerous online resources available for you to explore. We strongly recommend that you take advantage of these resources to deepen your understanding of ggplot(). 

Knit the document into a PDF and submit it to Canvas by 02/28/2024 at 11:59 pm.

# Load Packages and Data

*******

```{r}
####
#00#
####
library(knitr)
library(tidyverse)
library(datasets)
library(ggplot2)

```

*******

# 1. Make a Scatter Plot With Customized Point Size and Transparency (3pts)

*******

Description and Interpretation:  
I used the data ToothGrowth available in R. It was study done in Guinea pigs to understand the effect of different vitamin C on the length of odontoblasts. I used geom_jitter to visualize odontoblasts length in two differnt Vit C sources and colorcoded the doses. 

```{r}
####
#01#
?ToothGrowth
head(ToothGrowth)
summary(ToothGrowth)

ToothGrowth %>%
  ggplot(aes(x =supp, y = len, color = dose)) + 
  geom_jitter(size = 5,
              alpha = 0.5,
              width =  0.2,
              height = 0) +
  labs(x = "Supplement", y = "Length",
       title = "Length of odontoblasts in Guinea Pigs under different Vitamin C supplement") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_color_gradient(low = "blue", high = "red")
####


```

*******

# 2. Fit a Line Through Your Scatter Plot From 1. (3pts)

*******

Description and Interpretation:  
I used geom_abline to fit a line in the data for odontoblasts length only for Vit C supplement. I used linear model to find the intercept applied for the line. 

```{r}
####
#02#
ToothGrowth_filtered <- ToothGrowth %>% 
  filter(supp == "VC")
model <- lm(len ~ dose, data = ToothGrowth_filtered)
intercept <- coef(model)[1]

  ggplot(ToothGrowth_filtered, aes(x = dose, y = len)) + 
  geom_jitter(size = 5,
              alpha = 0.5,
              width =  0.2,
              height = 0) +
  geom_abline(intercept, slope = lm(len ~ dose, data = ToothGrowth_filtered)$coefficients[2], color = "blue") +
  labs(x = "Dose of VC", y = "Length",
       title = "Length of odontoblasts in Guinea Pigs under Vitamin C supplement") + 
  theme(plot.title = element_text(hjust = 0.5))
####
#your code here

```

# 3. Set the transparency, shape, and color of your graph from 1 based on expressions. Scale the color. (3pts)

*******

Description and Interpretation:  
I adjusted the transparency, shape from graph 1 but kept the same color gradient. 

```{r}
####
#03#
ToothGrowth %>%
  ggplot(aes(x =supp, y = len, color = dose)) + 
  geom_jitter(size = 5,
              alpha = 0.6,
              width =  0.2,
              height = 0,
              shape = 19) +
  labs(x = "Supplement", y = "Length",
       title = "Length of odontoblasts in Guinea Pigs under different Vitamin C supplement",
       subtitle = "OJ = Orange Juice VC = Vit. C") + 
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) +
  scale_color_gradient(low = "blue", high = "red")
####
#your code here

```

*******

# 4. Make a Boxplot With Customized Boxplot Width (3pts)

*******

Description and Interpretation:  
I made a boxplot with customized width of 0.4. The graph is for length of odontoblasts under differnt vit C supplement

```{r}
####
#04#
ToothGrowth %>%
  ggplot(aes(x =supp, y = len)) +
  geom_boxplot(width = 0.4) +
  labs(x = "Supplement", y = "Length",
       title = "Length of odontoblasts in Guinea Pigs under different Vitamin C supplement",
       subtitle = "OJ = Orange Juice VC = Vit. C") + 
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))

####
#your code here

```

*******

# 5. Overlay the Individual Points Over Your Box Plot From 3. and Adjust the Point Size and Transparency as Needed (3pts)

******

Description and Interpretation:  
The graph for this question is basically a combination of boxplot and jitter. I kept the color gradient same as before for dose. 

```{r}
####
#05#
ToothGrowth %>%
  ggplot(aes(x =supp, y = len, color = dose)) + 
  geom_boxplot(width = 0.4) +
  geom_jitter(size = 5,
              alpha = 0.6,
              width =  0.2,
              height = 0,
              shape = 19) +
  labs(x = "Supplement", y = "Length",
       title = "Length of odontoblasts in Guinea Pigs under different Vitamin C supplement",
       subtitle = "OJ = Orange Juice VC = Vit. C") + 
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) +
  scale_color_gradient(low = "blue", high = "red")

####
#your code here

```

# 6. Make a plot and modify the scale in a beneficial way using scale_x_continuous(), scale_y_continuous(), and coordinate_cartesian(). Explain how your modifications to the scales improved the plot. (4pts)

******

Description and Interpretation:  
For this question, I made a graph for odontoblasts length at different doses of supplement and separated supplements with colors. I think the scales made the graphs cleaner and tighter without unnecessary spaces around the edges. 

```{r}
####
#06#
#filter(ToothGrowth, supp == "VC") %>%
ToothGrowth %>%
  ggplot(aes(x =dose, y = len, color = supp)) + 
  geom_jitter(size = 5,
              alpha = 0.6,
              width =  0.2,
              height = 0,
              shape = 19) +
  labs(x = "Doses", y = "Length",
       title = "Length of odontoblasts in Guinea Pigs under different Vitamin C supplements") + 
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) +
    scale_x_continuous(breaks=c(0, 0.5, 1.0, 1.5, 2.0, 2.5), expand=c(0,0)) +
  scale_y_continuous(breaks=c(0, 5, 10, 15, 20, 25, 30, 35, 40), expand=c(0,0)) +
  coord_cartesian(xlim=c(0, 2.25), ylim=c(0, 35)) +
  scale_color_manual(breaks=c("OJ", "VC"), values=c("red", "green")) 
  
####
#your code here

```

******

# 7. Make a New (don't use any plots from 1-4) Plot with facet_wrap() (4pts)

******

Description and Interpretation:  
I made graphs for odontoblasts length at different doses for two separate supplements. 

```{r}
#07#
####
ToothGrowth %>%
  ggplot(aes(x =dose, y = len)) + 
  geom_jitter(size = 5,
              alpha = 0.6,
              width =  0.2,
              height = 0,
              color = "blue", shape = 19) +
  labs(x = "Doses", y = "Length",
       title = "Length of odontoblasts in Guinea Pigs under different Vitamin C supplements") + 
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) +
    scale_x_continuous(breaks=c(0, 0.5, 1.0, 1.5, 2.0, 2.5), expand=c(0,0)) +
  scale_y_continuous(breaks=c(0, 5, 10, 15, 20, 25, 30, 35, 40), expand=c(0,0)) +
  coord_cartesian(xlim=c(0, 2.25), ylim=c(0, 35)) +
  facet_wrap(~supp) 
#your code here

```

# 8. Using the plot from 5, scale the colors and reorder your facet to reverse order. (4pts)

******

Description and Interpretation:  
I reversed the order for facets using facet_wrap(~fct_rev(supp)).

```{r}
#08#
ToothGrowth %>%
  ggplot(aes(x =dose, y = len)) + 
  geom_jitter(size = 5,
              alpha = 0.6,
              width =  0.2,
              height = 0,
              shape = 19) +
  labs(x = "Doses", y = "Length",
       title = "Length of odontoblasts in Guinea Pigs under different Vitamin C supplements") + 
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) +
    scale_x_continuous(breaks=c(0, 0.5, 1.0, 1.5, 2.0, 2.5), expand=c(0,0)) +
  scale_y_continuous(breaks=c(0, 5, 10, 15, 20, 25, 30, 35, 40), expand=c(0,0)) +
  coord_cartesian(xlim=c(0, 2.25), ylim=c(0, 35)) +
  facet_wrap(~fct_rev(supp)) 
####
#your code here

```

******

# 9. Make a Plot Using a Geometry That Was Not Covered in Class (4pts)

******

Description and Interpretation:  
I used geom_violin and geom_area. I did not think they are very suitable for this data type.

```{r}
####
#09#
ToothGrowth %>%
  ggplot(aes(x =dose, y = len)) + 
  geom_violin(size = 1,
              color = "blue", fill ="lightblue") +
  labs(x = "Doses", y = "Length",
       title = "Length of odontoblasts in Guinea Pigs under different Vitamin C supplements") + 
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) +
    scale_x_continuous(breaks=c(0, 0.5, 1.0, 1.5, 2.0, 2.5), expand=c(0,0)) +
  scale_y_continuous(breaks=c(0, 5, 10, 15, 20, 25, 30, 35, 40), expand=c(0,0)) +
  coord_cartesian(xlim=c(0, 2.25), ylim=c(0, 35)) +
  facet_wrap(~supp) 

ToothGrowth %>%
  ggplot(aes(x =dose, y = len)) + 
  geom_area(size = 1,
              color = "blue", fill ="lightblue") +
  labs(x = "Doses", y = "Length",
       title = "Length of odontoblasts in Guinea Pigs under different Vitamin C supplements") + 
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) +
    scale_x_continuous(breaks=c(0, 0.5, 1.0, 1.5, 2.0, 2.5), expand=c(0,0)) +
  scale_y_continuous(breaks=c(0, 5, 10, 15, 20, 25, 30, 35, 40), expand=c(0,0)) +
  coord_cartesian(xlim=c(0, 2.25), ylim=c(0, 35)) +
  facet_wrap(~supp)
####
#your code here 
```

# 10. Make a Plot Using a ggplot Extension That Was Not Covered in Class. Explain what the extension does. (4pts)

******

Description and Interpretation:  
I tried to use gganimate but I could not get the output. It created the images for graphs frame-by-frame but did not create the animation. I tried to troubleshoot but I could not figure out the solution. I would appreciate if you could give me feedback on this. 

```{r}
####
#10#
library(gganimate)
library(gifski)

 animation1 <-  ToothGrowth %>%
  ggplot(aes(x =dose, y = len, color = supp)) + 
  geom_jitter(size = 5,
              alpha = 0.6,
              width =  0.2,
              height = 0,
              shape = 19) +
  labs(x = "Doses", y = "Length",
       title = "Length of odontoblasts in Guinea Pigs under different Vitamin C supplements") + 
  theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5)) +
    scale_x_continuous(breaks=c(0, 0.5, 1.0, 1.5, 2.0, 2.5), expand=c(0,0)) +
  scale_y_continuous(breaks=c(0, 5, 10, 15, 20, 25, 30, 35, 40), expand=c(0,0)) +
  coord_cartesian(xlim=c(0, 2.25), ylim=c(0, 35)) +
  scale_color_manual(breaks=c("OJ", "VC"), values=c("red", "green")) +
  transition_states(supp, transition_length = 2, state_length = 1) +
  enter_fade() +
    exit_shrink()
 animate(animation1, duration = 5, fps = 20, width = 200, height = 200, renderer = gifski_renderer())
anim_save("animation.gif", animation1 )

####
#your code here 

```

******

# Appendix

```{r show-code, ref.label = all_labels(), echo = TRUE, eval = FALSE}

```



